Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots

### 2023-01-18 sensors-2114505.tex
for final review

% Authors, for the paper (add full first names)
\Author{\hl{Francesco} %MDPI: Please carefully check the accuracy of names and affiliations.
 Roscia $^{1}$, Andrea Cumerlotti $^{1,2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi $^{1,3,}$*}

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Dynamic Legged Systems (DLS) Lab, Istituto Italiano di Tecnologia (IIT), \hl{Genoa,} %MDPI: Please add the postal code (or ZIP code in the U.S.). If the postal code is not available, please provide the P.O. Box. The following highlights are the same.
 Italy; \linebreak  \hl{francesco.roscia@iit.it} %MDPI: We added all the emails in affiliations according to the submitting system. Please confirm.
 (F.R.); andrea.cumerlotti@studenti.unitn.it (A.C.); claudio.semini@iit.it (C.S.)\\
$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, \hl{Trento}, Italy; andrea.delprete@unitn.it (A.D.P.)\\
$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, \hl{Trento}, Italy\\
}

% Contact information of the corresponding author
\corres{Correspondence: michele.focchi@unitn.it}

@@

\MF{This work introduces} a compact OCS, based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}. \MF{The open-source hardware and software are reported in \cite{github_solo}. Figure \ref{fig:solo12flywheels} shows the OCS mounted on the trunk of the robot.} The contributions of the paper are the following:
\begin{itemize}
	\item the design of a novel OCS that enables effective control of the orientation of a legged robot during a jump, while keeping the design simple. 
	In particular, the axes of rotation of the flywheels are set to be incident, enabling \hl{\textit{continuous}} %MDPI: Please confirm if all the italics in the maintext should be retained.
 controllability in both directions of roll and pitch, while keeping the device compact.
	\item simulations  with the quadruped Solo12 were conducted and demonstrated the effectiveness of the proposed approach.
\end{itemize}

\begin{figure}[H]%[t!] 1

	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
	\caption{\small The proposed Orientation Control System for 
		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes. Each wheel is located in a 3D-printed shell and mounted on the trunk of the body.}
	\label{fig:solo12flywheels}
\end{figure}

@@

\section{Background}
\label{sec:background}
The starting point for any OCS is the Euler's equation. For any mechanical system, the time derivative of the angular \hl{momentum}, %MDPI: Please confirm if all the bold in equations should be retained.
 $\bm{L}$ computed in respect to a reference point $O$, fixed in an inertial frame, equals the sum of the moments $\bm{M}_i$ applied to the system with respect to the same reference point:
\begin{equation}
\dot{\bm{L}} = \sum_i \bm{M}_i
\end{equation}
\hl{When} %MDPI: Please confirm if the noindent should be retained. The following highlights are the same.
 the result of the external moments applied to the system is zero, and the Euler's equation simplifies to:
 \begin{equation}
\dot{\bm{L}} = 0 \quad \Rightarrow \quad \bm{L}(t) = const,
\label{eq:cons_ang_mom}
\end{equation}
which is known as conservation of angular momentum.
Referring to legged robots, this condition occurs when the system is not in contact with the ground or with other objects, e.g., during a fall or the flight phase of a jump.
In this case, it is possible to change the angular velocity of the base link, thereby changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{wieber2016modeling}. If the angular momentum of a certain body changes, then one of the others must change to maintain the total sum as constant.
\subsection*{Preliminary Analysis}
Most quadruped robots are designed in such a way that the largest amount of the mass is located in the main body. As a consequence, the contribution of the moments due to the acceleration of other bodies (i.e., leg links) is moderate. \\
As an example, consider the case of Solo12. Even if each limb accounts for about $13.4\%$ of the total mass, changes in the total inertia of the robot come with the motion of only the upper leg (thigh) and the lower leg (calf), since the displacement of distribution of the hip mass with the system's CoM is approximately constant. \MF{In view of this, each limb is responsible for only $7.5\%$ of the total mass in varying the total inertia of the robot}. Flywheels can be used to alleviate the lack of control authority.

@@

	\includegraphics[width=.7\linewidth]{figures/elroys-beanie.eps}
	\caption{\small Schematic representation of the Elroy's Beanie model used for the preliminary analysis of the pitch motion.}
	\label{fig:elroy-beanie}
\end{figure}

This consists of two rigid bodies connected through their CoMs with a revolute joint. One of the bodies represents the robot in its nominal configuration and the other models the OCS, here represented by a single wheel for the sake of simplicity. The aim was to examine the rotational dynamics of the system as a whole. To simplify the analysis, in the following \MF{only the effects on the pitch angle of the robot was considered}, keeping in mind that the same arguments also apply to the roll. 
Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system's CoM.
The angular momentum $L$ of this system can be written as:
\begin{equation}\label{eq:ang_mom_elroy}
L(t) = \left(I_r + I_{f}\right)\dot{\theta}(t) + I_{f}\dot{\gamma}(t)
\end{equation}
where $\dot{\theta}$ and $\dot{\gamma}$ are, respectively, the robot;s pitch rate and the angular speed of the wheel.
\MF{The robot is driven to reach} a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$, given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
\begin{equation}
I_{f} \geq I_r \dfrac{ \Delta \dot{\theta} }{\Delta \dot{\theta}+\dot{\gamma}_{max}}.
\end{equation}
where $\Delta \dot{\theta} =  \left\lvert \dot{\theta}_0-\dot{\theta}_{des} \right\rvert$ is the base velocity variation.
Figure \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.

@@

	\caption{\MF{Top view of Solo12 with the proposed OCS. The $x-$ and $y-$axes embody the forward and left directions of the robot base, respectively. The figure also illustrates the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$  that represent the axes of rotation of the left flywheel and of the right one and $\alpha$ is the angle of incidence, defined in Section \ref{subseq:flywheel_axes}.}}
	\label{fig:axes}
\end{figure}
\vspace{-11pt}
\begin{figure}[H]%[t]5
 \hspace{-16pt}
	\includegraphics[width=1\linewidth]{figures/inertiaEllipsoid.eps}
	\vspace{-9pt}
	\caption{\hl{The} %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 inertia tensor can be seen as an ellipsoid. Its principal axes are in the direction of the eigenvectors of the tensor and their lengtha depend on the eigenvalues.}
	\label{fig:ellipsoid}
\end{figure}
In the case of Solo12, this angle was about $40^{\circ}$. With these considerations in mind, the angular momentum produced by the wheels, expressed in the robot base frame, was:
\begin{equation}
\begin{split}
\tensor[_b]{\bm{L}}{_f} &= \tensor[_b]{\bm{L}}{_{fl}} + \tensor[_b]{\bm{L}}{_{fr}} \\
&=I_{f, zz} \, \bm{\omega}_{fl} + I_{f, zz} \, \bm{\omega}_{fr}\\
&=I_{f, zz}
\left[\begin{array}{c}
\left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha) \\ 
\left(\omega_{fl} +  \omega_{fr}\right)\cos(\alpha) \\
0
\end{array}\right]
\end{split}
\label{eq:bLf}
\end{equation}
in which $\bm{\omega}_{fl}= \omega_{fl} \, \hat{\bm{a}}_{l}$ and $\bm{\omega}_{fr}= \omega_{fr} \, \hat{\bm{a}}_{r}$ are the angular velocity vectors of the two wheels, and $\omega_{fl}$ and $\omega_{fr}$ are the angular speeds provided to each flywheel by its actuation system. The latter equation shows that the difference of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Figure \ref{fig:sum-diff}. Using the definition of $\bm{C}$, \eqref{eq:bLf} is rewritten as:
\begin{equation}
\begin{split}
\tensor[_b]{\bm{L}}{_f} & = I_{f,zz}
\left[
\begin{array}{cc}
\sin(\alpha) & -\sin(\alpha) \\
\cos(\alpha) & \cos(\alpha) \\
0 & 0
\end{array}
\right]
\left[
\begin{array}{cc}
\omega_{fl} \\
\omega_{fr}
\end{array}
\right] \\
&= I_{f,zz} \bm{C} \bm{\omega}_f.
\end{split}
\end{equation}
\vspace{-6pt}
\begin{figure}[H]%6

@@

\subsection{Inertia Selection}
Once the desired inertia $I_f$ is selected, according to the 2D Elroy's Beanie model, it is used to realize the 3D OCS (Figure \ref{fig:flywheel}). 
\MF{Two identical flywheels were designed to have} the shape of hollow cylinders. This shape increases the inertia by locating the mass far away from the rotation axis. The inertia tensor expressed in its principal axes, \linebreak  $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders' inner and outer radii, $r$ and $R$, its height $h$ and material density $\rho$:
\begin{equation}
\begin{split}
I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
\end{split}
\end{equation}
\hl{Notice} that in the Elroy's Beanie model, there is a single body that models the complete OCS. The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$. 
The parameters $R$ and $h$ can be set to have a compact OCS and $\rho$ depends on the chosen material, which was stainless steel in our case. The inner radius $r$ could be adjusted to obtain the desired inertia. Spokes with negligible mass connected the wheel to the motor shaft. All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.

@@

\vspace{-6pt}
\begin{table}[H]%[t] 1
	\caption{\hl{Sizes} %MDPI: Please confirm if all the "\cdot" between numbers should be a multiplication sign ("\times" U+00D7).
 and dynamic parameters of a single flywheel.}
	\label{tab:dim_fw}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CCC}
		\toprule
		\textbf{Parameter} & \textbf{Value} & \textbf{Unit}\\
		\midrule
		$r$ & $2.20 \cdot 10^{-2}$ & $\mathrm{m}$ \\
		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$\rho$   & $7.86 \cdot 10^{+3}$ & $\mathrm{kg/m^3}$ \\
		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
		\bottomrule	\end{tabularx}
\end{table}

\subsection{Flywheels' Control Law}
\label{subsec:control_law}
To derive a control law based on the robot's base orientation, we made use of \eqref{eq:cons_ang_mom}, expressing all the contributions to the time derivative of total angular momentum with respect to the base reference frame:
\begin{equation}
\tensor[_b]{\bm{I}}{_f} \, \tensor[_b]{\dot{\bm{\omega}}}{_f} + \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\dot{\bm{\omega}}}{_r} + \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right) = 0
\end{equation}
\hl{From} this expression, the moment on the base caused by the acceleration of the flywheels \MF{is defined}, and can be used as a feedback torque $\bm{\tau}_{fb}$:
\begin{equation}
\begin{split}
\bm{\tau}_{fb} &= \bm{I}_f \dot{\bm{\omega}}_f \\
&= - \tensor[_b]{\bm{I}}{_r} \dot{\bm{\omega}}_b - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right)
\end{split}
\end{equation}
\hl{Proportional} and derivative action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$ \MF{is used}, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive definite gain matrices for the error in attitude and angular velocity.
The orientation error $\bm{e} \in SO(3)$ requires that the algebra of the special rotational group is computed. To avoid singular configurations, orientations \MF{are represented} with quaternions.
The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocities of the base:
\begin{equation}
\bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right).
\end{equation}
\hl{Projecting} the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, the control action $\bm{u}$ \MF{yields}
\begin{equation}
\label{eq:ctrl_law}
\bm{u} = \bm{C}^T \bm{\tau}_{fb}.
\end{equation}

\section{Results of Simulations}
\label{sec:simulations}
To validate our OCS, three simulations on different scenarios {are reported.~The authors} wanted to test the capability of the proposed approach in terms of the following: 
to reject a disturbance when the robot was in the flight phase of a jump, and to dampen trunk oscillations after touch-down, 
and to work in parallel with the joints of the legs to achieve a highly dynamic motion. 
All the simulations were performed using \MF{ROS \cite{quigley2009ros}, together with Gazebo \cite{koenig2004design}. The physics-related parameters are reported in Table \ref{tab:pys_param}.} References for the joints of the legs were computed off-line using Crocoddyl, an optimal control library for robots based on Differential Dynamic Programming (DDP) algorithms. It uses Pinocchio for fast computation of robot dynamics and the analytical derivatives. References $\bm{q}_{ref}$, $\dot{\bm{q}}_{ref}$ and $\bm{\tau}_{ref}$ for joint positions, velocities and torques were then executed on-line with a proportional-derivative joint controller:
\begin{equation}
\bm{\tau}_j = \bm{K}_{p,\, j} (\bm{q}_{ref} - \bm{q}) + \bm{K}_{d,\, j} (\dot{\bm{q}}_{ref} - \dot{\bm{q}}) + \bm{\tau}_{ref}
\end{equation}
\hl{The} OCS was commanded to track base orientation references using the control law introduced in Section \ref{subsec:control_law} \MF{coded in Python}. ({\hl{The} %MDPI: Footnote is not permitted in our journal. We include this paragraph to the maintext. Please confirm.
 video, with all the simulations, is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}  (\hl{accessed} %MDPI: Please add the access date (Format: Date Month Year). e.g., (accessed on 1 January 2020).
 on).)}

@@

\bibitem[Barasuol \em{et~al.}(2013)Barasuol, Buchli, Semini, Frigerio,
  De~Pieri, and Caldwell]{barasuol2013reactive}
Barasuol, V.; Buchli, J.; Semini, C.; Frigerio, M.; De~Pieri, E.R.; Caldwell,
  D.G.
\newblock A reactive controller framework for quadrupedal locomotion on
  challenging terrain.
\newblock In Proceedings of the 2013 IEEE International Conference on Robotics
  and Automation, \hl{Karlsruhe, Germany, 6--10 May 2013;} %MDPI: Newly added information. Please confirm. The following highlights are the same.
 IEEE: \hl{Piscataway, NJ, USA,}  2013; pp.~2554--2561.

>>>>

% Authors, for the paper (add full first names)
\Author{Francesco %MDPI: Please carefully check the accuracy of names and affiliations.
 Roscia $^{1}$, Andrea Cumerlotti $^{1,2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi $^{1,3,}$*}

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Dynamic Legged Systems (DLS) Lab, Istituto Italiano di Tecnologia (IIT), 16163 Genova, %MDPI: Please add the postal code (or ZIP code in the U.S.). If the postal code is not available, please provide the P.O. Box. The following highlights are the same.
 Italy; \linebreak francesco.roscia@iit.it %MDPI: We added all the emails in affiliations according to the submitting system. Please confirm.
 (F.R.); claudio.semini@iit.it (C.S.)\\
$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, 38123 Trento, Italy; andrea.cumerlotti@studenti.unitn.it (A.C.); andrea.delprete@unitn.it (A.D.P.)\\
$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, 38123 Trento, Italy\\
}

% Contact information of the corresponding author
\corres{Correspondence: michele.focchi@unitn.it (M.F.)}

@@

\MF{This work introduces} a compact OCS, based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}. \MF{The open-source hardware and software are reported in \cite{github_solo}. Figure \ref{fig:solo12flywheels} shows the OCS mounted on the trunk of the robot.} The contributions of the paper are the following:
\begin{itemize}
	\item the design of a novel OCS that enables effective control of the orientation of a legged robot during a jump, while keeping the design simple. 
	In particular, the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} %MDPI: Please confirm if all the italics in the maintext should be retained.
 controllability in both directions of roll and pitch, while keeping the device compact.
	\item simulations  with the quadruped Solo12 were conducted and demonstrated the effectiveness of the proposed approach.
\end{itemize}

\begin{figure}[H]%[t!] 1
	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
	\caption{\small The proposed Orientation Control System for 
		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes. Each wheel is located in a 3D-printed shell and mounted on the trunk of the body.}
	\label{fig:solo12flywheels}
\end{figure}

@@

\section{Background}
\label{sec:background}
The starting point for any OCS is the Euler's equation. For any mechanical system, the time derivative of the angular momentum, %MDPI: Please confirm if all the bold in equations should be retained.
 $\bm{L}$ computed in respect to a reference point $O$, fixed in an inertial frame, equals the sum of the moments $\bm{M}_i$ applied to the system with respect to the same reference point:
\begin{equation}
\dot{\bm{L}} = \sum_i \bm{M}_i
\end{equation}
When %MDPI: Please confirm if the noindent should be retained. The following highlights are the same.
 the result of the external moments applied to the system is zero, and the Euler's equation simplifies to:
 \begin{equation}
\dot{\bm{L}} = 0 \quad \Rightarrow \quad \bm{L}(t) = const,
\label{eq:cons_ang_mom}
\end{equation}
which is known as conservation of angular momentum.
Referring to legged robots, this condition occurs when the system is not in contact with the ground or with other objects, e.g., during a fall or the flight phase of a jump.
In this case, it is possible to change the angular velocity of the base link, thereby changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{wieber2016modeling}. If the angular momentum of a certain body changes, then one of the others must change to maintain the total sum as constant.
\subsection*{Preliminary Analysis}
Most quadruped robots are designed in such a way that the largest amount of the mass is located in the main body. As a consequence, the contribution of the moments due to the acceleration of other bodies (i.e., leg links) is moderate. \\
As an example, consider the case of Solo12. Even if each limb accounts for about $13.4\%$ of the total mass, changes in the total inertia of the robot come with the motion of only the upper leg (thigh) and the lower leg (calf), since the displacement of distribution of the hip mass with the system's CoM is approximately constant. \MF{In view of this, each limb is responsible for only $7.5\%$ of the total mass in varying the total inertia of the robot}. Flywheels can be used to alleviate the lack of control authority.

@@

	\includegraphics[width=.7\linewidth]{figures/elroys-beanie.eps}
	\caption{\small Schematic representation of the Elroy's Beanie model used for the preliminary analysis of the pitch motion.}
	\label{fig:elroy-beanie}
\end{figure} 
\noindent This consists of two rigid bodies connected through their CoMs with a revolute joint. One of the bodies represents the robot in its nominal configuration and the other models the OCS, here represented by a single wheel for the sake of simplicity. The aim was to examine the rotational dynamics of the system as a whole. To simplify the analysis, in the following \MF{only the effects on the pitch angle of the robot was considered}, keeping in mind that the same arguments also apply to the roll. 
Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system's CoM.
The angular momentum $L$ of this system can be written as:
\begin{equation}\label{eq:ang_mom_elroy}
L(t) = \left(I_r + I_{f}\right)\dot{\theta}(t) + I_{f}\dot{\gamma}(t)
\end{equation}
where $\dot{\theta}$ and $\dot{\gamma}$ are, respectively, the robot;s pitch rate and the angular speed of the wheel.
\MF{The robot is driven to reach} a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$, given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
\begin{equation}
I_{f} \geq I_r \dfrac{ \Delta \dot{\theta} }{\Delta \dot{\theta}+\dot{\gamma}_{max}}.
\end{equation}
where $\Delta \dot{\theta} =  \left\lvert \dot{\theta}_0-\dot{\theta}_{des} \right\rvert$ is the base velocity variation.
Figure \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.

@@

	\caption{\MF{Top view of Solo12 with the proposed OCS. The $x-$ and $y-$axes embody the forward and left directions of the robot base, respectively. The figure also illustrates the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$  that represent the axes of rotation of the left flywheel and of the right one and $\alpha$ is the angle of incidence, defined in Section \ref{subseq:flywheel_axes}.}}
	\label{fig:axes}
\end{figure}
\vspace{-11pt}
\begin{figure}[H]%[t]5
 \hspace{-16pt}
	\includegraphics[width=1\linewidth]{figures/inertiaEllipsoid.eps}
	\vspace{-9pt}
	\caption{The %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 inertia tensor can be seen as an ellipsoid. Its principal axes are in the direction of the eigenvectors of the tensor and their lengths depend on the eigenvalues.}
	\label{fig:ellipsoid}
\end{figure}
In the case of Solo12, this angle was about $40^{\circ}$. With these considerations in mind, the angular momentum produced by the wheels, expressed in the robot base frame, was:
\begin{equation}
\begin{split}
\tensor[_b]{\bm{L}}{_f} &= \tensor[_b]{\bm{L}}{_{fl}} + \tensor[_b]{\bm{L}}{_{fr}} \\
&=I_{f, zz} \, \bm{\omega}_{fl} + I_{f, zz} \, \bm{\omega}_{fr}\\
&=I_{f, zz}
\left[\begin{array}{c}
\left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha) \\ 
\left(\omega_{fl} +  \omega_{fr}\right)\cos(\alpha) \\
0
\end{array}\right]
\end{split}
\label{eq:bLf}
\end{equation}
in which $\bm{\omega}_{fl}= \omega_{fl} \, \hat{\bm{a}}_{l}$ and $\bm{\omega}_{fr}= \omega_{fr} \, \hat{\bm{a}}_{r}$ are the angular velocity vectors of the two wheels, and $\omega_{fl}$ and $\omega_{fr}$ are the angular speeds provided to each flywheel by its actuation system. The latter equation shows that the difference of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Figure \ref{fig:sum-diff}. Using the definition of $\bm{C}$, \eqref{eq:bLf} is rewritten as:
\begin{equation}
\begin{split}
\tensor[_b]{\bm{L}}{_f} & = I_{f,zz}
\left[
\begin{array}{cc}
\sin(\alpha) & -\sin(\alpha) \\
\cos(\alpha) & \cos(\alpha) \\
0 & 0
\end{array}
\right]
\left[
\begin{array}{cc}
\omega_{fl} \\
\omega_{fr}
\end{array}
\right] \\
&= I_{f,zz} \bm{C} \bm{\omega}_f.
\end{split}
\end{equation}
\vspace{-6pt}
\begin{figure}[H]%6

@@

\subsection{Inertia Selection}
Once the desired inertia $I_f$ is selected, according to the 2D Elroy's Beanie model, it is used to realize the 3D OCS (Figure \ref{fig:flywheel}). 
\MF{Two identical flywheels were designed to have} the shape of hollow cylinders. This shape increases the inertia by locating the mass far away from the rotation axis. The inertia tensor expressed in its principal axes, \linebreak  $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders' inner and outer radii, $r$ and $R$, its height $h$ and material density $\rho$:
\begin{equation}
\begin{split}
I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
\end{split}
\end{equation}
Notice that in the Elroy's Beanie model, there is a single body that models the complete OCS. The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$. 
The parameters $R$ and $h$ can be set to have a compact OCS and $\rho$ depends on the chosen material, which was stainless steel in our case. The inner radius $r$ could be adjusted to obtain the desired inertia. Spokes with negligible mass connected the wheel to the motor shaft. All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.

@@

\vspace{-6pt}
\begin{table}[H]%[t] 1
	\caption{Sizes %MDPI: Please confirm if all the "\cdot" between numbers should be a multiplication sign ("\times" U+00D7).
 and dynamic parameters of a single flywheel.}
	\label{tab:dim_fw}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CCC}
		\toprule
		\textbf{Parameter} & \textbf{Value} & \textbf{Unit}\\
		\midrule
		$r$ & $2.20 \times 10^{-2}$ & $\mathrm{m}$ \\
		$R$ & $3.00\times 10^{-2}$ & $\mathrm{m}$ \\
		$h$   & $1.02\times 10^{-2}$ & $\mathrm{m}$ \\
		$\rho$   & $7.86 \times 10^{+3}$ & $\mathrm{kg/m^3}$ \\
		$m$   & $1.02\times 10^{-1}$ & $\mathrm{kg}$ \\
		$I_{xx}, I_{yy}$   & $3.64 \times 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
		$I_{zz}$   & $7.11 \times 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
		\bottomrule	\end{tabularx}
\end{table}

\subsection{Flywheels' Control Law}
\label{subsec:control_law}
To derive a control law based on the robot's base orientation, we made use of \eqref{eq:cons_ang_mom}, expressing all the contributions to the time derivative of total angular momentum with respect to the base reference frame:
\begin{equation}
\tensor[_b]{\bm{I}}{_f} \, \tensor[_b]{\dot{\bm{\omega}}}{_f} + \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\dot{\bm{\omega}}}{_r} + \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right) = 0
\end{equation}
From this expression, the moment on the base caused by the acceleration of the flywheels \MF{is defined}, and can be used as a feedback torque $\bm{\tau}_{fb}$:
\begin{equation}
\begin{split}
\bm{\tau}_{fb} &= \bm{I}_f \dot{\bm{\omega}}_f \\
&= - \tensor[_b]{\bm{I}}{_r} \dot{\bm{\omega}}_b - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right)
\end{split}
\end{equation}
Proportional and derivative action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$ \MF{is used}, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive definite gain matrices for the error in attitude and angular velocity.
The orientation error $\bm{e} \in SO(3)$ requires that the algebra of the special rotational group is computed. To avoid singular configurations, orientations \MF{are represented} with quaternions.
The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocities of the base:
\begin{equation}
\bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right).
\end{equation}
Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, the control action $\bm{u}$ \MF{yields}
\begin{equation}
\label{eq:ctrl_law}
\bm{u} = \bm{C}^T \bm{\tau}_{fb}.
\end{equation}

\section{Results of Simulations}
\label{sec:simulations}
To validate our OCS, three simulations on different scenarios {are reported.~The authors} wanted to test the capability of the proposed approach in terms of the following: 
to reject a disturbance when the robot was in the flight phase of a jump, and to dampen trunk oscillations after touch-down, 
and to work in parallel with the joints of the legs to achieve a highly dynamic motion. 
All the simulations were performed using \MF{ROS \cite{quigley2009ros}, together with Gazebo \cite{koenig2004design}. The physics-related parameters are reported in Table \ref{tab:pys_param}.} References for the joints of the legs were computed off-line using Crocoddyl, an optimal control library for robots based on Differential Dynamic Programming (DDP) algorithms. It uses Pinocchio for fast computation of robot dynamics and the analytical derivatives. References $\bm{q}_{ref}$, $\dot{\bm{q}}_{ref}$ and $\bm{\tau}_{ref}$ for joint positions, velocities and torques were then executed on-line with a proportional-derivative joint controller:
\begin{equation}
\bm{\tau}_j = \bm{K}_{p,\, j} (\bm{q}_{ref} - \bm{q}) + \bm{K}_{d,\, j} (\dot{\bm{q}}_{ref} - \dot{\bm{q}}) + \bm{\tau}_{ref}
\end{equation}
The OCS was commanded to track base orientation references using the control law introduced in Section \ref{subsec:control_law} \MF{coded in Python}. (The %MDPI: Footnote is not permitted in our journal. We include this paragraph to the maintext. Please confirm.
video, with all the simulations, is available at web link \url{https://www.dropbox.com/s/6py2avxj2f9kksn/video.mp4?dl=0}, accessed %MDPI: Please add the access date (Format: Date Month Year). e.g., (accessed on 1 January 2020).
 on 18 January 2023).

@@

\bibitem[Barasuol \em{et~al.}(2013)Barasuol, Buchli, Semini, Frigerio,
  De~Pieri, and Caldwell]{barasuol2013reactive}
Barasuol, V.; Buchli, J.; Semini, C.; Frigerio, M.; De~Pieri, E.R.; Caldwell,
  D.G.
\newblock A reactive controller framework for quadrupedal locomotion on
  challenging terrain.
\newblock In Proceedings of the 2013 IEEE International Conference on Robotics
  and Automation, Karlsruhe, Germany, 6--10 May 2013; %MDPI: Newly added information. Please confirm. The following highlights are the same.
 IEEE: Piscataway, NJ, USA,  2013; pp.~2554--2561.
